---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# kwb.en13508.2

[![Appveyor build status](https://ci.appveyor.com/api/projects/status/i5xx4npr86rg783h/branch/master?svg=true)](https://ci.appveyor.com/project/KWB-R/kwb-en13508-2/branch/master)
[![Build Status](https://travis-ci.org/KWB-R/kwb.en13508.2.svg?branch=master)](https://travis-ci.org/KWB-R/kwb.en13508.2)
[![codecov](https://codecov.io/github/KWB-R/kwb.en13508.2/branch/master/graphs/badge.svg)](https://codecov.io/github/KWB-R/kwb.en13508.2)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/kwb.en13508.2)](http://cran.r-project.org/package=kwb.en13508.2)

This R package allows you to read and write data on sewerage inspections in
the text format described in the European Norm 
[EN 13508-2](http://www.dwa.de/dwa/shop/shop.nsf/Produktanzeige?openform&produktid=P-DWAA-8KTG5R).

## Install the Package

You can install the latest development version of the package from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("KWB-R/kwb.en13508.2", build_vignettes = TRUE)
```

## Load the Package

```{r}
library(kwb.en13508.2)
```

## Read a File Encoded in EN 13508.2 Format

You can read a file formatted in EN 13508.2 text format using 
`readEuCodedFile()`. 

For the purpose of demonstration, we load an example file that is contained 
in the package:

```{r}
# Set the relative path to the example file
file_path <- "extdata/example_13508_2.txt"

# Set the absolute path to the example file in the package folder
input_file <- system.file(file_path, package = "kwb.en13508.2")

# Read the example file
survey <- readEuCodedFile(input_file)
```

The result is a list with the three components representing the A-, B- and C-
parts, respectively, of the file:

* `header.info`: A-part containing information on the file format
* `inspections`: B-parts containing information on the inspections as a whole
* `observations`: C-parts containing information on the detailed observations 
that were made during the inspections

### Part A: Information on the file format

```{r}
survey$header.info
```

### Part B: Inspections

```{r}
head(survey$inspections)
```

The codes used as column names correspond to the codes defined in the norm. 
We provide a function `getCodes()` in the package that returns a table mapping 
these codes to their meanings:

```{r}
# Get all codes and meanings defined in the European Norm
code_info <- getCodes(fields = c("Code", "Text_EN"))

# Show the meanings for the codes that are used in the table of inspections
code_info[code_info$Code %in% names(survey$inspections), ]
```


### Part C: Observations

```{r}
head(survey$observations)
```

Again, let's have a look at what the columns mean:

```{r}
# Show the meanings for the codes that are used in the table of inspections
code_info[code_info$Code %in% names(survey$observations), ]
```

The column `inspno` groups together observations that belong to the same 
inspection. The number in the column refers to the row number in the table of 
inspections (`survey$inspections`). For example, to get all the observations 
that belong to the third inspection, you may filter `survey$observations` like 
in the following. For a more compact output we exclude the tenth column ("F")
containing the remarks:

```{r}
survey$observations[survey$observations$inspno == 3, -10]
```

The information on the inspection is found in the `inspno`-th row 
(here: third row) of `config$inspections`:

```{r}
survey$inspections[3, ]
```

## Write a File Encoded in EN 13508.2 Format

Once you have prepared a list with the three components `header.info`, 
`inspections` and `observations` as described above, you can use the function
`writeEuCodedFile()` to write a file formatted in EN 13508.2-format:

```{r}
# Define the path to an output file
output_file <- file.path(tempdir(), "example_en13508.2.txt")

# Write the CCTV survey data to the file
writeEuCodedFile(survey, output_file)
```
The first 20 lines of the file produced read:

```{r}
kwb.utils::catLines(readLines(output_file, 20))
```

